Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib

### 2023-09-20 texor.Rmd
phew! finally have a flow diagram. Layout with diagrammer is an absolute mess and takes a lot of work to get it looking ok

1. Articles can include interactive graphics and tables. 
2. The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers. (See @W3C for current guidelines.)
3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
4. Search engines can easily access the full text of articles, facilitating discovery of published articles.

<!-- traditional HTML is the most universally accessible format - Australian Human Rights Commission https://humanrights.gov.au/our-work/disability-rights/world-wide-web-access-disability-discrimination-act-advisory-notes-ver and https://printdisability.org -->

The latter points are motivations for converting all of the legacy R Journal articles into HTML.

@@

```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'}
#
if (knitr::is_html_output()) {
  library(DiagrammeR)
  DiagrammeR::mermaid(diagram = "flowchart TD
    A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
    B --> |pandoc reader|C{AST} 
    C -->|pandoc writer| D[Converted Article]
    C -->|modify AST using lua filter| C{AST}")
}
if (knitr::is_latex_output()) {
  knitr::include_graphics("workflow.png")
}
```

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element. In an R Markdown code chunk generating a figure, this can be set using `fig.alt`. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

>>>>

1. Articles can include interactive graphics (to allow readers to query, re-focus, watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information). 
2. It is the best format for accessibility [@Daisy]. Screen readers and other assistive technologies can deliver the content to print disabled readers. (See @W3C for current guidelines.)
3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
4. Search engines can easily access the full text of articles, facilitating discovery of published articles.

These provide the motivations for converting all of the legacy R Journal articles into HTML.

@@

```{r workflow, fig.cap="Workflow of the document conversion conducted by texor. (Note: AST is pandoc's abstract syntax tree.)", out.width="100%", fig.align='center', fig.width=4, fig.height=3, layout="l-body", fig.alt="Flow diagram showing original article pre-processed in R passed to pandoc which creates the AST and modifies using many lua filters then passed to pacdoc writer to produce the converted article."}
library(DiagrammeR)
wf <- create_node_df(
        n=8,
        type=c("a", "b", "a", "b", 
               "a", "b", "b", "a"),
        label=c("original article", "pre-processing in R",
                "pre-processed article", "pandoc reader",
                "AST", 
                "modify AST using lua filter",
                "pandoc writer", 
                "converted article"
                ),
        shape="rectangle",
        width=c(1.3, 1.7, 1.9, 1.3, 0.5, 2.3, 1.2, 1.5),
        fillcolor = c("darkseagreen2", "grey85", "darkseagreen2", "grey85", "darkseagreen2", "grey85", "grey85", "darkseagreen2"),
        color = c("black", "grey85", "black", "grey85", "black", "grey85", "grey85", "black"),
        fontcolor="black",
        x=c(1,1,1,1,1,2,1,1),
        y=c(1,2,3,4,5,5,6,6),
        fontsize=14
        )
wf_arrows <- create_edge_df(from = c(1, 2, 3, 4, 5, 6, 5, 7),
                            to = c(2, 3, 4, 5, 6, 5, 7, 8),
                            color="black")
wf_graph <- create_graph(wf, wf_arrows, 
                   directed = TRUE, attr_theme = "tb")
set.seed(7)
render_graph(wf_graph, layout="fr", height=500)
```

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element. In an R Markdown code chunk generating a figure, this can be set using `fig.alt`. Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress: to encourage adoption by authors of new article submissions, and it would be useful to edit into the converted legacy articles at a later date.

### 2023-09-20 texor.tex
phew! finally have a flow diagram. Layout with diagrammer is an absolute mess and takes a lot of work to get it looking ok

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Articles can include interactive graphics and tables.
\item
  The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers. (See Sheri Byrne-Haber (2023) for current guidelines.)
\item
  HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
\item
  Search engines can easily access the full text of articles, facilitating discovery of published articles.
\end{enumerate}

The latter points are motivations for converting all of the legacy R Journal articles into HTML.

@@

{\centering \includegraphics[width=0.9\linewidth]{workflow}

@@

\caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow}
\end{figure}

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element. In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

@@

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Articles can include interactive graphics (to allow readers to query, re-focus, watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information).
\item
  It is the best format for accessibility (The Daisy Consortium 2023). Screen readers and other assistive technologies can deliver the content to print disabled readers. (See Sheri Byrne-Haber (2023) for current guidelines.)
\item
  HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
\item
  Search engines can easily access the full text of articles, facilitating discovery of published articles.
\end{enumerate}

These provide the motivations for converting all of the legacy R Journal articles into HTML.

@@

{\centering \includegraphics[width=1\linewidth]{texor_files/figure-latex/workflow-1}

@@

\caption{Workflow of the document conversion conducted by texor. (Note: AST is pandoc's abstract syntax tree.)}\label{fig:workflow}
\end{figure}

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element. In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}. Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress: to encourage adoption by authors of new article submissions, and it would be useful to edit into the converted legacy articles at a later date.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Daisy}{}}%
The Daisy Consortium. 2023. {``Creating the Best Way to Read and Publish.''} \url{https://daisy.org}.
